Exercise 3: Software Architecture in Practice Case: Beer Web Store

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Adding performance quality attribute

>>>>

As we have only defined the Availability and Performance requirements, only these will be dealt with here, yet naturally requirements concerning the other quality attributes must also be handled.

\subsubsection{Availability}
Availability has two main concerns; finding out that a resource is unavailable and if it is, then restart or switch to another resource.

Finding out if a resource is down may be done using one of the fault detection tactice; ping/echo, heartbeat or exception. Due to the loose coupling between the client (browser) and the server, the exception solution is not benificial. Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g. 1 second to a maintance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintanance service can act.

The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}. As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and tehn if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds. As 30 seconds is insufficient to bott a webserver and a database, the Spare tactic would be insufficient. At the same tiem having Active redundancy would be over-kill, as a this kind of immidiat switch-over is not required. For these reasons Passive Redundancy is sufficient.

As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server. The user may experience beeing thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability. During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable. for this it will be benificial to use the \emph{Rollback} tactic on the \emph{Basket} component.

Finally a web-server may be deployed multiple times on teh same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic. This is e.g. done by the Apache-server.

As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database. However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
Depending on the binding between teh database and the server, it may also be possibel to use the exceptions tactic between teh server and the database, so the server itself can restart the database, if an exception occur in a call to the database.

\subsubsection{Performance}

As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve propper performance.

Due to the lowing price of hardware and the increasing price of developement time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resourcez}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cach the entire result in preperation for a search refinement. If the database is remotely located the caching of results becomes even more important.

There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under developement control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconcivably expensive compare to simply increasing the computational resources.
